---
title: "R_Func_Data_Rotate"
author: "Lihe Liu"
date: "11/9/2019"
output:
  html_document: 
    highlight: textmate
    theme: readable
---

# User Instructions  

### Descriptions  

This function performs (partial) data pivotal-transformation. For each row (here, **each minute in a day**), the selected <span style="color:blue">*columns names*</span> and the corresponding <span style="color:red">*entry values*</span> (here, <span style="color:blue">**animal ID**</span> and the corresponding <span style="color:red">**actions counts**</span>)  will be re-positioned to new columns.  

For example, here, for each **each minute in a day**, the selected <span style="color:blue">**animal ID**</span> the corresponding <span style="color:red">**actions counts**</span> will be re-positioned to new columns - <span style="color:blue">**"Animal"**</span> and <span style="color:red">**"Action"**</span>. Please note, the mintue time in the **unselected** column will be duplicated.   

This gives us what we want --- 

**"In each minute in a day, we track the active animal and their action counts"**

### Inputs  
Users are required to provide the original dataset, **Date** and **Minutes** are in two seperated (or one) columns. Note,  

  - first column *must* be date in formation
  - date information formate **must** be "xx/xx/xx"
  - keyword: Next, provide the keyword of your output (e.g. "Farm_1_rotate") suffix will append by itself then you will get "Farm_1_rotate.txt" or "Farm_1_rotate.RData", depends on argument "output_type"
  - output_type: can be ".txt" (*default*) or ".RData"
  
### Features  

  - <span style="color:blue">**Does not rely on pre-defined position of header row**</span>, no need to count which rows has the animal id information. we differentiate "Date" from "xx/xx/xx" by detection of the presense of "/" (**so, be sure to have "/" in the date entries, as least once!**)
  - <span style="color:blue">**Does not rely on balance records**</span>, no need to have equal number of records every day. We detect "data sections" by using the postion of header rows, so however many rows in the 
middle will be extracted. It works like setting flags (like the SNPs position on the chip), each time we extract the records between two flags.

```{r functino pre}
Pivot_Rotate = function(test, # Input data, must have the same structure as you have
                        safe = "/", # to differentiate "mm/dd/yy"
                        # Next, provide the keyword of your output (e.g. "Farm_1_rotate")
                        # suffix will append by itself
                        # then you will get "Farm_1_rotate.txt" or "Farm_1_rotate.RData", 
                        #depends on argument "output_type"
                        keyword = "", 
                        output_type = "txt"){# can be ".txt" or ".RData"
  library(tidyverse)
  All_names = names(sort(table(test[,1]))) #Obtain all names to dertermin all possible header names
  detect_header_index = str_extract(All_names, "/")# Use "/" to detect: differentiate "mm/dd/yy" from c("Date","Day",..), etc
  seperate_name = All_names[is.na(detect_header_index)] # Obtain values for exclusion
  
  #Next step - get header_index using matching, the famous %in%, this much safer than count by HANDS!!!
  header_index = which(test[,1] %in% seperate_name)
  # create a container for output
  final = data.frame(Date = character(),
                     Minutes = character(),
                     Animal = character(),
                     Action = numeric())
  # loop along all the header_row and parse the section(from head to tail)
  for (i in seq_along(header_index)){
    # We are using one bin point (i) and the next bin point (i+1) at the same time
    # HOWEVER, this only works when (i < total bin points)
    # because when (i == total), (i+1) will be out of bound!
    #
    # Situation 1: when ( i < total )
    if ( i < length(header_index)){ 
      target = header_index[i] # aim at the sepcific header row
      header_row_tmp = as.vector(test[target,-c(1:2)]) # take the actually cow id in the headerrow as the column names of the following sub-dataset (time section)
      header_row = as.character(header_row_tmp[!is.na(header_row_tmp)]) # remove NAs
      section_data = test[c((target + 1):(header_index[i+1]-1)),c(1:(length(header_row)+2))] # select section data using two bins (like head and tail)
    } else { # while (i == total), we use the total number of rows as the "tail"
      section_data = test[c((target + 1):nrow(test)),c(1:(length(header_row)+2))]
    }
    colnames(section_data) = c("Date","Minutes",header_row) 
    # perform roate using external function
    rotate_tpm <- section_data %>% 
      pivot_longer(header_row, names_to = "Animal", values_to = "Action") %>% 
      drop_na() # remove na, just in case
      # potentially could customize column names when needed. uncomment these and renamme
      # %>% rename(New_name_1 =Date,
      #            New_name_1 = Minutes,
      #            New_name_1= animal,
      #            New_name_1 = action)
    final = rbind(final,rotate_tpm)
    message("Now working on day ", i,"/",length(header_index), "---",section_data[1,1])
  }
  if (output_typ == "txt"){
    write.table(final,file = paste(keyword,".txt",sep = ""),
                quote = F,row.names=FALSE)
  } else {save(final, file = paste(keyword,".RData",sep = ""))}
}
```


```{r testing1 }
####==================####
#==     Farm1        ===#
#####===============####
#Farm_1_raw = read.table("Victoria.txt",header = F, sep = "\t")
#Pivot_Rotate(test = Farm_1_raw,
#             safe = "/",
#             keyword = "Farm_1_Rotated")
```



```{r check1 }
### For check with the original

#library(tidyverse)
#load("Farm_1_Rotated.RData")
#final %>% dplyr::filter(Date == "7/1/2018",
#                        Minutes == "0:06") %>%  print()
```


```{r testing 2}
####==================####
#==     Farm2        ===#
###==================### 
#Together_raw = read.table("Together_Jacob.txt",header = F, sep = "\t")
#Pivot_Rotate(test = Together_raw,
#             safe = "/",
#             keyword = "Together_Jacob_Rotated")
```

